submission

### 2024-03-09 submission.Rmd
feat: ex1.3 power.t.test we need to figure this one out

```{r,warning=FALSE}
library(readr)
library(dplyr)
library(tidyverse)
library(MASS)
library(mixtools)
```

@@

```{r}
wilcox.test(diabetic_adult$Glu, diabetic_seniors$Glu, alternative="two.sided")
```

**TODO :: Figure out this error...**

@@

# Combine the data frames
combined_data <- rbind(diabetic_active, diabetic_inactive)
rmarkdown::paged_table(combined_data)
```

Immediately we can observe that we have significantly less data than for the other exercises, which can indicate that test results can be inconclusive.

@@

@@

## 2.2  We now wish to perform the previous test for all of the 3116 genes. For each of the genes, fit a linear model that explains `cholesterol` as a function of the gene expression and compute the p-value of the test. AND

## 2.3  Order the p-values and plot the ordered p-values as a function of their rank. On the same plot, display the line $y = x/3116$. Discuss the result.

@@

```
Plotting all values, we can conclude that there are amount of models that have a p-value below 0.05. It is tempting to make the conclusion that all these models are statistically significant since they all have a small p-value. However, in the event that the we only have random data, we still expect 5\% of the models to appear significantly significant. In fact, if we were to only have random data, we expect a uniform distribution between 0 and 1 for the p-values. Visually, looking at the Figure above, we can the p-values perform better that the uniform distribution and therefore appear to contain significant findings.

@@

```
 In the nest question, we are tasked with finding the maximum number of genes such that the expected number of false discoveries is at most 5\%. For this method, we use the Benjamini-Hochberg procedure \cite{FDR}. In short, it states that if we have a set of $\{ H_1, H_2, ... \: , H_m \}$ with associated and ordered p-values $\{ p_1, p_2, ... \: , p_m \}$, then the threshold for the index can be written as:

@@

```
We can see that $\alpha \frac{i}{m}$ plots out a line, and we choose i such that $P_{(i)}$ always is smaller or equal to that value. Doing so, we get 941 discoveries such that the expected number of false discoveries is no more than 5 \%

##2.5 A stronger guarantee would be that on impose that there shouldn't be more than a 5 \% risk that any discovery is false. For this method, we use the Holm–Bonferroni method. The method is as follows:\\

>>>>

```{r,warning=FALSE}
library(readr)
library(dplyr)
library(tidyverse)
library(MASS)
library(mixtools)
library(pwr)
library(mclust)
```

@@

```{r}
delta_wilcox <- wilcox.test(diabetic_seniors$Glu, diabetic_adult$Glu, alternative="two.sided")
delta_wilcox
```

Based on these results we can observe that we cannot reject the null-hypothesis, thus, we can say that there is not enough evidence to suggest that the mean blood glucose between diabetic adults and diabetic seniors is different.

@@

# Combine the data frames
combined_data <- rbind(diabetic_active, diabetic_inactive)
rmarkdown::paged_table(combined_data)

ggplot(combined_data, aes(x = Group, y = Glu)) +
  geom_boxplot() +
  labs(x = "Activity Group", y = "Glucose") +
  ggtitle("Glucose Distribution for Diabetics by Activity") +
  theme_minimal()

ggplot(combined_data, aes(x = Group, y = Glu, fill = Group)) +
  geom_violin(trim = FALSE) +
  labs(x = "Activity Group", y = "Glucose") +
  ggtitle("Glucose Distribution for Diabetics by Activity") +
  theme_minimal()
```

Immediately we can observe that we have significantly less data than for the other exercises, which can indicate that test results can be inconclusive. Furthermore, we cannot necessarily make an assumption about the normality of the data due to the fact that we have such little samples. It could for example be the case that for many more samples that the distribution looks much more like a normal.

However, for now we will not make any normality assumption and we will be using a Wilcoxon test where we will analyse its power of varying number of samples for $\alpha=0.05$ which will enable us to see how the power of the test evolves over the number of samples.

```{r}
delta_wilcox  <- wilcox.test(diabetic_active$Glu, diabetic_inactive$Glu, alternative="two.sided")$statistic
alpha <- 0.05
seq_n <- seq(10,1)
power_wilcox  <- map_dbl(seq_n, ~power.t.test(n = .x, delta_wilcox , sig.level = alpha)$power)
data.frame(power = c(power_wilcox), n = rep(seq_n),
           method = rep(c("Wilcox"), each = length(seq_n))) %>% 
  ggplot() + aes(x = n, y = power, color = method) + geom_line()
```

We observe that for a sample size of 4, the power of the Wilcoxon t-test is more or less 1.0 meaning that it is almost guaranteed that the null-hypothesis will be rejected when it is false. Furthermore, we plot for $n=4$ the different values of $\alpha$ wrt. the power of the Wilcoxon test.

```{r}
delta_wilcox  <- wilcox.test(diabetic_active$Glu, diabetic_inactive$Glu, alternative="two.sided")$statistic
seq_alpha <- 10^seq(-12, -5, .1)
n <- 4
power_wilcox  <- map_dbl(seq_alpha, ~power.t.test(n, delta_wilcox , sig.level = .x)$power)
data.frame(power = c(power_wilcox), alpha = rep(seq_alpha, 1),
           method = rep(c("Wilcox"), each = length(seq_alpha))) %>% 
  ggplot() + aes(x = alpha, y = power, color = method) + geom_line()
```

As observed earlier for a $\alpha=0.05$ we basically have guaranteed that the null-hypothesis is rejected correctly only after 2 samples. However, this seems strange and thus one might want to adjust the $\alpha$ value to get a more reasonable sample requirement.

@@

```{r}
diabete_and_nondiabete = filter(data, Diabete != 3)
contingency_table <- table(diabete_and_nondiabete$Sex, diabete_and_nondiabete$Diabete)
contingency_table
```

```{r}
fisher.test(contingency_table)
```

```{r}
diabetes = c(8, 13)
population = c(nrow(filter(data, Sex==1)), nrow(filter(data, Sex==2)))

prop.test(diabetes, population, correct = FALSE)
```

```{r}
prop.test(diabetes, population, correct = TRUE)
```

@@

### 2.2 We now wish to perform the previous test for all of the 3116 genes. For each of the genes, fit a linear model that explains `cholesterol` as a function of the gene expression and compute the p-value of the test. AND

### 2.3 Order the p-values and plot the ordered p-values as a function of their rank. On the same plot, display the line $y = x/3116$. Discuss the result.

@@

Plotting all values, we can conclude that there are amount of models that have a p-value below 0.05. It is tempting to make the conclusion that all these models are statistically significant since they all have a small p-value. However, in the event that the we only have random data, we still expect 5% of the models to appear significantly significant. In fact, if we were to only have random data, we expect a uniform distribution between 0 and 1 for the p-values. Visually, looking at the Figure above, we can the p-values perform better that the uniform distribution and therefore appear to contain significant findings.

@@

In the nest question, we are tasked with finding the maximum number of genes such that the expected number of false discoveries is at most 5%. For this method, we use the Benjamini-Hochberg procedure \cite{FDR}. In short, it states that if we have a set of $\{ H_1, H_2, ... \: , H_m \}$ with associated and ordered p-values $\{ p_1, p_2, ... \: , p_m \}$, then the threshold for the index can be written as:

@@

We can see that $\alpha \frac{i}{m}$ plots out a line, and we choose i such that $P_{(i)}$ always is smaller or equal to that value. Doing so, we get 941 discoveries such that the expected number of false discoveries is no more than 5 %

##2.5 A stronger guarantee would be that on impose that there shouldn't be more than a 5 % risk that any discovery is false. For this method, we use the Holm–Bonferroni method. The method is as follows:\\
